The inefficacy of superficial similarities for improving instructor-student relationships

### 2021-04-26 outputs/paper/paper.Rmd
update repo organization, edit paper.

---
title: "The inefficacy of superficial similarities for improving instructor-student relationships"
author: "Amy Farrow"
date: "April 19th, 2021"
header-includes:
  \usepackage{dcolumn}
  \usepackage{float}
  \usepackage{rotating}
  \usepackage{pdflscape}
  \usepackage{booktabs}
output:
  bookdown::pdf_document2:
    toc: yes
subtitle: "Replicating ‘Taking It to the Next Level’"
abstract: "This paper replicates the 2019 article 'Taking It to the Next Level' [@citeoriginal], which evaluates an intervention to improve college instructor-student relationships. The modeling results indicate that the intervention, consisting of informing instructors and students about commonalities, has a weak positive effect on student perceptions of instructor-student similarity, but no effect on student perceptions of instructor-student relationship, instructor perception of similarity, or instructor-student relationship. The goal of this intervention was to improve student performance and persistence, and while the scalability and affordability are desirable, there are no measurable effects on any of the student outcome measures, including grades and reenrollment. These results are consistent with the original paper.\\par \\textbf{Keywords:} replication study, instructor-student relationships, college, similarity, college persistence"
thanks: 'Code and data are available at: [github.com/amycfarrow/takingittothenextlevelrepro](https://github.com/amycfarrow/takingittothenextlevelrepro).'
bibliography: references.bib
---

@@

library(tidyverse)
library(bookdown)    # for cross referencing figures and graphs; referencing
library(kableExtra)  # for nicer tables
library(here) # for working in an RProject
library(Hmisc) # for correlation matrix
library(readstata13) # for stata files
library(finalfit) # for summary table
library(stargazer) # for model table
library(MASS) # for ordinal logistic model
library(cowplot) # for putting plots side by side
library(ltm) # for checking scale validity
library(DiagrammeR) # for causal diagram
library(olsrr)
library(tidymodels) # for train test

@@

```{r, warning=FALSE, message=FALSE, echo=FALSE}
# The original coding of race contained errors -- ex. a professor who selected White and Asian in the survey,
# and was marked as Asian for most cases, but white for one random case.
# This remakes the variable t_race, for consistency
expdata <- expdata %>%
  mutate(t1_race_1 = replace_na(as.numeric(t1_race_1), 0),
         t1_race_2 = replace_na(as.numeric(t1_race_2), 0),
         t1_race_3 = replace_na(as.numeric(t1_race_3), 0),
         t1_race_4 = replace_na(as.numeric(t1_race_4), 0),
         t1_race_5 = replace_na(as.numeric(t1_race_5), 0),
         t1_race_6 = replace_na(as.numeric(t1_race_6), 0),
         t1_race_7 = replace_na(as.numeric(t1_race_7), 0)) %>%
  mutate(t_race = ifelse(t1_race_1 + t1_race_2 + t1_race_3 + t1_race_4 + t1_race_5 + t1_race_6 + t1_race_7 > 1, 
                         "Mixed race", 
                         ifelse(t1_race_1 + t1_race_2 + t1_race_3 + t1_race_4 + t1_race_5 + t1_race_6 + t1_race_7 < 1, "Unknown",
                                ifelse(t1_race_1 == 1, "White",
                                       ifelse(t1_race_2 == 1, "Black",
                                              ifelse(t1_race_3 == 1, "Hispanic",
                                                     ifelse(t1_race_4, "Asian",
                                                            ifelse(t1_race_5, "American Indian",
                                                                   ifelse(t1_race_6, "Middle Eastern",
                                                                          "Other")))))))))

expdata <- expdata %>%
  mutate(s1_race_1 = replace_na(as.numeric(s1_race_1), 0),
         s1_race_2 = replace_na(as.numeric(s1_race_2), 0),
         s1_race_3 = replace_na(as.numeric(s1_race_3), 0),
         s1_race_4 = replace_na(as.numeric(s1_race_4), 0),
         s1_race_5 = replace_na(as.numeric(s1_race_5), 0),
         s1_race_6 = replace_na(as.numeric(s1_race_6), 0),
         s1_race_7 = replace_na(as.numeric(s1_race_7), 0)) %>%
  mutate(s_race = ifelse(s1_race_1 + s1_race_2 + s1_race_3 + s1_race_4 + s1_race_5 + s1_race_6 + s1_race_7 > 1,
                         "Mixed race", 
                         ifelse(s1_race_1 + s1_race_2 + s1_race_3 + s1_race_4 + s1_race_5 + s1_race_6 + s1_race_7 < 1, "Unknown",
                                ifelse(s1_race_1 == 1, "White",
                                       ifelse(s1_race_2 == 1, "Black",
                                              ifelse(s1_race_3 == 1, "Hispanic",
                                                     ifelse(s1_race_4, "Asian",
                                                            ifelse(s1_race_5, "American Indian",
                                                                   ifelse(s1_race_6, "Middle Eastern",
                                                                          "Other"))))))))) %>%
  mutate(s_hisp_black = ifelse((s1_race_2 == 1 | s1_race_3 == 1), 1, 0))

@@

This paper uses the R statistical language [@citeR] and `tidyverse` packages [@citetidyverse]. `here` was used to manage the RProject [@citehere]. Stata files were handled using `readstata13` [@citereadstata13]. The report was made using RMarkdown [@citermarkdown] and `bookdown` [@citebookdown]. Tables were made using `kableExtra` [@citekableExtra], `finalfit` [@citefinalfit], and `stargazer` [@citestargazer]. The correlation matrix was made using functions by @apacorrcode using the `Hmisc` package [@citeHmisc]. `MASS` provided additional models [@citeMASS]. Plots used `cowplot` [@citecowplot]. `ltm` was used to check scale validity [@citeltm]. Finally, the diagram was made using `DiagrammeR` [@citeDiagrammeR].

@@

Table \@ref(tab:demotable1) shows the student covariates for the treatment and control samples. There are no significant differences between the treatment and control groups.

```{r demotable1, echo=FALSE, warning=FALSE, message=FALSE}
# Make a summary table of nominal variables, by treatment group.
demotab <- expdata %>% 
  dplyr::select(treatment, 
                s_female, 
                s_race, 
                s_firstgen,
                s1_age,
                ir_f16_gpa,
                year) %>%
  mutate(year = as.factor(year)) %>%
  mutate(treatment = ff_label(treatment, "Group"),
         s_female = ff_label(s_female, "Student gender"), 
         s1_age = ff_label(s1_age, "Student age"),
         s_race = ff_label(s_race, "Student race"), 
         ir_f16_gpa = ff_label(ir_f16_gpa, "CPGA"),
         s_firstgen = ff_label(s_firstgen, "Student first-gen status"), 
         year = ff_label(year, "Year")) %>%
  summary_factorlist("treatment", 
                     c("s_female", "s_race", "s_firstgen", "year", "s1_age", "ir_f16_gpa"),
                     p = TRUE, 
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_col_totals = TRUE,
                     add_row_totals = TRUE,
                     include_row_missing_col = TRUE,
                     col_totals_rowname = "",
                     total_col = TRUE,
                     col_totals_prefix = "N(%) = ") %>%
  rename(N = "Total N", Missing = "Missing N")

@@

```{r demotable2, echo=FALSE, warning=FALSE, message=FALSE}
# make a summary table of continuous variables
demotab2 <- expdata %>%
  dplyr::select(teacherid, t_age, t_firstgen, t_race, t_female, n_course) %>%
  unique() %>%
  mutate(teacherid = ifelse(as.numeric(teacherid) > 50, "0", "1")) %>%
  mutate(t_age = ff_label(t_age, "Instructor age"), 
         t_female = ff_label(t_female, "Instructor gender"),
         t_race = ff_label(t_race, "Instructor race"),
         t_firstgen = ff_label(t_firstgen, "Instructor first-gen status"),
         n_course = ff_label(n_course, "Course size")) %>%
  summary_factorlist("teacherid",
                     c("t_female", "t_race", "t_firstgen", "t_age", "n_course"),
                     p = TRUE, 
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_row_totals = TRUE,
                     include_row_missing_col = TRUE,
                     col_totals_rowname = "",
                     total_col = TRUE,
                     add_col_totals = TRUE,
                     col_totals_prefix = "N = ") %>%
  rename(N = "Total N", Missing = "Missing N")

@@

```{r racegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(8,4), fig.cap="Bar chart illustrating racial identification proprtions of participants"}
instructorrace <- expdata %>%
  dplyr::select(teacherid,
         t1_race_1, t1_race_2, t1_race_3, t1_race_4, t1_race_5, t1_race_6, t1_race_7, t_multi_race) %>%
  unique() %>%
  dplyr::summarize(White = sum(t1_race_1, na.rm = TRUE)/n(),
            Black = sum(t1_race_2, na.rm = TRUE)/n(),
            Hispanic = sum(t1_race_3, na.rm = TRUE)/n(),
            Asian = sum(t1_race_4, na.rm = TRUE)/n(),
            "American Indian" = sum(t1_race_5, na.rm = TRUE)/n(),
            "Middle Eastern" = sum(t1_race_6, na.rm = TRUE)/n(),
            Other = sum(t1_race_7, na.rm = TRUE)/n(),
            Multiple = sum(t_multi_race > 1, na.rm = TRUE)/n())
studentrace <- expdata %>%
  dplyr::select(id,
         s1_race_1, s1_race_2, s1_race_3, s1_race_4, s1_race_5, s1_race_6, s1_race_7, s_multi_race) %>%
  unique() %>%
  dplyr::summarize(White = sum(s1_race_1, na.rm = TRUE)/n(),
            Black = sum(s1_race_2, na.rm = TRUE)/n(),
            Hispanic = sum(s1_race_3, na.rm = TRUE)/n(),
            Asian = sum(s1_race_4, na.rm = TRUE)/n(),
            "American Indian" = sum(s1_race_5, na.rm = TRUE)/n(),
            "Middle Eastern" = sum(s1_race_6, na.rm = TRUE)/n(),
            Other = sum(s1_race_7, na.rm = TRUE)/n(),
            Multiple = sum(s_multi_race > 1, na.rm = TRUE)/n())

bind_cols(as_tibble(cbind(nms = names(instructorrace), t(instructorrace))),
          as_tibble(cbind(nms = names(studentrace), t(studentrace))))%>%
  dplyr::select(-nms...3) %>%
  pivot_longer(!nms...1, names_to = "group", values_to = "portion") %>%
  mutate(nms...1 = fct_relevel(nms...1, 
                               c("Hispanic", "White", "Asian",  "Middle Eastern",
                                 "Black", "American Indian", "Other", "Multiple")),
         group = fct_relevel(group, c("V2...4", "V2...2"))) %>%
  ggplot() +
  geom_col(aes(x = nms...1, y = as.numeric(portion), fill = group), position = "dodge") +
  labs(x = "Self-identified race", y = "Proportion of sample", 
       fill = "Group", title = "Self-identified racial proportions differ for instructors and students") +
  scale_fill_manual(values = c("#c8b5f7", "#f9b5c9"), labels = c("Students", "Instructors")) +
  theme_minimal()
```

@@

```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(9,3), fig.cap="Histogram illustrating class sizes", out.width="100%", fig.align="center"}
n_instructor <- expdata %>%
  dplyr::select(teacherid, n_course) %>%
  unique()

@@

title <- ggdraw() + 
  draw_label(
    "Some classes are very large (means marked)",
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, 0, 7)
  )

plot_grid(title,
          combined,
          legend,
          ncol = 1,
          rel_heights = c(.1, 2))
```

@@

Table \@ref(tab:summarytable) shows summary statistics for key variables identified by @citeoriginal, including missing values. There do not appear to be significant differences between the treatment and control groups, other than in student similarity perception. Even in these two measures, the difference appears to be small (mean scores of 3.4 versus 3.6 and 3.5 versus 3.6).

@@

$X_{1i}$ is a vector of student-level covariates (student ISR anticipation, gender, CGPA, year of study). ISR anticipation is used as a control for end-of-term student ISR perception, while gender and CGPA are used as controls for grade based outcomes. @citeoriginal state this is because females earn higher grades generally.

$X_{2j}$ is a vector of instructor-level covariates (class size).

@@

Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception (t2_sim1).

@@

Equation \@ref(eq:logmodel) is @citeoriginal's logistic model, used for the binary outcome, which is enrollment in Fall term 2017 (f17_enrolled).

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Make all replication models
studentfirstsimilaritymodel <- lm(data = expdata,
                                  formula = s1_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
                                     na.action = na.omit)
studentsecondsimilaritymodel <- lm(data = expdata,
                                   formula = s2_sim ~ treatment + n_course + s_female + ir_f16_gpa + year,
                                     na.action = na.omit)
studentsecondrelationshipmodel <- lm(data = expdata,
                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s_female + ir_f16_gpa + year,
                                     na.action = na.omit)
teachersecondrelationshipmodel <- lm(data = expdata,
                                     formula = t2_tsr ~ treatment + n_course + s_female + ir_f16_gpa + year,
                                     na.action = na.omit)
grademodel <- lm(data = expdata,
                 formula = grade ~ treatment + n_course + s_female + ir_f16_gpa + year,
                                     na.action = na.omit)
examgrademodel <- lm(data = expdata %>% filter(obj_exam == 1),
                     formula = t2_finalexam ~ treatment + n_course + s_female + ir_f16_gpa + year,
                                     na.action = na.omit)
teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s_female + ir_f16_gpa + year,
                                data = expdata,
                                Hess = TRUE,
                                     na.action = na.omit)
enrollmentfallmodel <- glm(data = expdata,
                           formula = f17_enrolled ~ treatment + n_course + s_female + ir_f16_gpa + year,
                           family = "binomial",
                                     na.action = na.omit)
```

In Table \@ref(tab:modeltable), Model 1 shows a statistically significant relationship between treatment and initial student similarity perception ($\beta$ = 0.194, 95% CI [0.136, 0.253]). Model 2 shows a statistically significant relationship between treatment and student end-of-term similarity perception ($\beta$ = 0.11, 95% CI [0.038, 0.183]). However, both coefficients are minimal on a scale of 1 to 5.

Model 3 shows no significant relationship between treatment and student end-of-term ISR perception. However, it does show a significant correlation between student ISR anticipation and student end-of-term ISR perception ($\beta$ = 0.671, 95% CI [0.624, 0.717]). For every 1 point increase in ISR anticipation on a 1 to 5 scale, the expected increase in end-of-term ISR perception is 0.671.

Models 4, 5, 6, 7, and 8 show no significant relationship between treatment and instructor end-of-term ISR perception, course grade, objectively graded exam grade, instructor perception of similarity at the end of the term, and enrollment in the subsequent term.

Models 1 through 8 were also repeated with additional student covariates (gender, race, first-generation status, CGPA, and year), and the results still showed a treatment effect on student similarity perception but no other effect.

Model validation for the two primary significant models, Models 1 and 2 from Table \@ref(tab:modeltable), can be found in [Appendix D].

```{r modeltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
# Make table to display model results
stargazer(studentfirstsimilaritymodel,
          studentsecondsimilaritymodel,
          studentsecondrelationshipmodel,
          teachersecondrelationshipmodel,
          grademodel,
          examgrademodel,
          teachersecondsimilarity,
          enrollmentfallmodel,
          title = "Replication model results",
          align = TRUE,
          ci = TRUE,
          float.env = "sidewaystable",
          notes.label = "",
          no.space = TRUE,
          keep.stat = c("n", "rsq", "adj.rsq", "logrank", "aic", "res.dev", "ser", "sigma2"),
          column.sep.width = "-10pt",
          label = "tab:modeltable",
          type = "latex",
          header = FALSE,
          initial.zero = FALSE,
          #omit = c("teacherid"),
          #omit.labels = c("Instructor fixed effect"),
          dep.var.labels = c("Student sim. 1",
                             "Student sim. 2",
                             "Student ISR 2",
                             "Instructor ISR 2",
                             "Course grade",
                             "Final grade",
                             "Instructor sim. 2",
                             "Persistence"),
          covariate.labels = c("Treatment", "Student ISR 1", "Course size", "Student female", "CGPA", "Year of study"))
```

@@

Similarly, the subset of first-generation college student participants is considered. While the effect of treatment on student similarity perception is larger than for the general student population and the Black and/or Hispanic student participants ($\beta$ = 0.265, 95% CI [0.177, 0.353] for initial perception; $\beta$ = 0.15, 95% CI [0.039, 0.26] for end-of-term perception), there is still no effect on ISR perception, instructor perception, or student outcomes. Full results are in [Appendix F].

>>>>

---
title: "The inefficacy of superficial similarities for improving instructor-student relationships"
author: "Amy Farrow"
date: "April 26th, 2021"
header-includes:
  \usepackage{dcolumn}
  \usepackage{float}
  \usepackage{rotating}
  \usepackage{pdflscape}
  \usepackage{booktabs}
output:
  bookdown::pdf_document2:
    toc: yes
subtitle: "Replicating ‘Taking It to the Next Level’"
abstract: "This paper replicates the 2019 article 'Taking It to the Next Level' [@citeoriginal], which evaluates an intervention to improve college instructor-student relationships. The modeling results indicate that the intervention, consisting of informing instructors and students about commonalities, has a weak positive effect on student perceptions of instructor-student similarity, but no effect on student perceptions of instructor-student relationship, instructor perception of similarity, or instructor-student relationship. The goal of this intervention was to improve student performance and persistence, and while the scalability and affordability are desirable, there are no measurable effects on any of the student outcome measures, including grades and reenrollment. These results are consistent with the original paper.\\par \\textbf{Keywords:} replication study, instructor-student relationships, college, similarity, college persistence"
thanks: 'Code and data are available at: [github.com/amycfarrow/takingittothenextlevelrepro](https://github.com/amycfarrow/takingittothenextlevelrepro).'
bibliography: references.bib
---

@@

library(tidyverse)
library(bookdown)    # for cross referencing figures and graphs; referencing
library(kableExtra)  # for nicer tables
library(here) # for working in an RProject
library(Hmisc) # for correlation matrix
library(readstata13) # for stata files
library(finalfit) # for summary table
library(stargazer) # for model table
library(MASS) # for ordinal logistic model
library(cowplot) # for putting plots side by side
library(ltm) # for checking scale validity
library(DiagrammeR) # for causal diagram
library(tidymodels) # for train test
library(performance) # for model performance evaluation

@@

```{r, warning=FALSE, message=FALSE, echo=FALSE}
# The original coding of race contained errors -- ex. a professor who selected White and Asian in the survey,
# and was marked as Asian for most cases, but white for one random case.
# This remakes the variable t_race, for consistency

expdata <- expdata %>%
  mutate(t1_race_1 = replace_na(as.numeric(t1_race_1), 0),
         t1_race_2 = replace_na(as.numeric(t1_race_2), 0),
         t1_race_3 = replace_na(as.numeric(t1_race_3), 0),
         t1_race_4 = replace_na(as.numeric(t1_race_4), 0),
         t1_race_5 = replace_na(as.numeric(t1_race_5), 0),
         t1_race_6 = replace_na(as.numeric(t1_race_6), 0),
         t1_race_7 = replace_na(as.numeric(t1_race_7), 0)) %>%
  mutate(t_race = ifelse(t1_race_1 + t1_race_2 + t1_race_3 + t1_race_4 + t1_race_5 + t1_race_6 + t1_race_7 > 1, 
                         "Mixed race/ethnicity", 
                         ifelse(t1_race_1 + t1_race_2 + t1_race_3 + t1_race_4 + t1_race_5 + t1_race_6 + t1_race_7 < 1, "Unknown",
                                ifelse(t1_race_1 == 1, "White/Caucasian",
                                       ifelse(t1_race_2 == 1, "Black or African American",
                                              ifelse(t1_race_3 == 1, "Hispanic American or Latino/a",
                                                     ifelse(t1_race_4, "Asian or Pacific Islander",
                                                            ifelse(t1_race_5, "American Indian or Alaskan",
                                                                   ifelse(t1_race_6, "Middle Eastern",
                                                                          "Other Race")))))))))

expdata <- expdata %>%
  mutate(s1_race_1 = replace_na(as.numeric(s1_race_1), 0),
         s1_race_2 = replace_na(as.numeric(s1_race_2), 0),
         s1_race_3 = replace_na(as.numeric(s1_race_3), 0),
         s1_race_4 = replace_na(as.numeric(s1_race_4), 0),
         s1_race_5 = replace_na(as.numeric(s1_race_5), 0),
         s1_race_6 = replace_na(as.numeric(s1_race_6), 0),
         s1_race_7 = replace_na(as.numeric(s1_race_7), 0)) %>%
  mutate(s_race = ifelse(s1_race_1 + s1_race_2 + s1_race_3 + s1_race_4 + s1_race_5 + s1_race_6 + s1_race_7 > 1,
                         "Mixed race/ethnicity", 
                         ifelse(s1_race_1 + s1_race_2 + s1_race_3 + s1_race_4 + s1_race_5 + s1_race_6 + s1_race_7 < 1, "Unknown",
                                ifelse(s1_race_1 == 1, "White/Caucasian",
                                       ifelse(s1_race_2 == 1, "Black or African American",
                                              ifelse(s1_race_3 == 1, "Hispanic American or Latino/a",
                                                     ifelse(s1_race_4, "Asian or Pacific Islander",
                                                            ifelse(s1_race_5, "American Indian or Alaskan",
                                                                   ifelse(s1_race_6, "Middle Eastern",
                                                                          "Other Race"))))))))) %>%
  mutate(s_hisp_black = ifelse((s1_race_2 == 1 | s1_race_3 == 1), 1, 0))

@@

This paper uses the R statistical language [@citeR] and `tidyverse` packages [@citetidyverse]. `here` was used to manage the RProject [@citehere]. Stata files were handled using `readstata13` [@citereadstata13]. The report was made using RMarkdown [@citermarkdown] and `bookdown` [@citebookdown]. Tables were made using `kableExtra` [@citekableExtra], `finalfit` [@citefinalfit], and `stargazer` [@citestargazer]. The correlation matrix was made using functions by @apacorrcode using the `Hmisc` package [@citeHmisc]. `MASS` provided additional models [@citeMASS]. Plots used `cowplot` [@citecowplot]. `ltm` was used to check scale validity [@citeltm]. `tidymodels` [@citetidymodels] and `performance` [@citeperformance] were used in assessing model validity and performance. Finally, the diagram was made using `DiagrammeR` [@citeDiagrammeR].

@@

Most key demographic traits come from the initial surveys given to students and instructors. For gender, they were asked to identify as male, female, or other, and given the option to enter text. For race/ethnicity, they were allowed to check any number of a range of options ('White/Caucasian', 'Black or African American', 'Hispanic American or Latino/a', 'Asian or Pacific Islander', 'American Indian or Alaskan', 'Middle Eastern', and 'Other Race'). They were also asked if they were the first generation in their family to attend college (no or yes). Finally, they were asked to enter their age in years. The student's year of study and cumulative CGPA as of the previous term and the class size came from the college's internal records.

Table \@ref(tab:demotable1) shows the student covariates for the treatment and control samples. There are no significant differences between the treatment and control groups, as indicated by p-values which are all above 0.4.

```{r demotable1, echo=FALSE, warning=FALSE, message=FALSE}
# Make a summary table of nominal variables, by treatment group.
demotab <- expdata %>% 
  dplyr::select(treatment, 
                s1_female, 
                s_race, 
                s_firstgen,
                s1_age,
                ir_f16_gpa,
                year) %>%
  mutate(year = as.factor(year)) %>%
  mutate(treatment = ff_label(treatment, "Group"),
         s1_female = ff_label(s1_female, "Student gender"), 
         s1_age = ff_label(s1_age, "Student age"),
         s_race = ff_label(s_race, "Student race/ethnicity"), 
         ir_f16_gpa = ff_label(ir_f16_gpa, "CPGA"),
         s_firstgen = ff_label(s_firstgen, "Student first-gen status"), 
         year = ff_label(year, "Year")) %>%
  summary_factorlist("treatment", 
                     c("s1_female", "s_race", "s_firstgen", "year", "s1_age", "ir_f16_gpa"),
                     p = TRUE, 
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_col_totals = TRUE,
                     add_row_totals = TRUE,
                     include_row_missing_col = TRUE,
                     col_totals_rowname = "",
                     total_col = TRUE,
                     col_totals_prefix = "N(%) = ") %>%
  rename(N = "Total N", Missing = "Missing N")

@@

```{r demotable2, echo=FALSE, warning=FALSE, message=FALSE}
# make a summary table of continuous variables
demotab2 <- expdata %>%
  dplyr::select(teacherid, t_age, t_firstgen, t_race, t_female, n_course) %>%
  unique() %>%
  mutate(teacherid = ifelse(as.numeric(teacherid) > 50, "0", "1")) %>%
  mutate(t_age = ff_label(t_age, "Instructor age"), 
         t_female = ff_label(t_female, "Instructor gender"),
         t_race = ff_label(t_race, "Instructor race/ethnicity"),
         t_firstgen = ff_label(t_firstgen, "Instructor first-gen status"),
         n_course = ff_label(n_course, "Class size")) %>%
  summary_factorlist("teacherid",
                     c("t_female", "t_race", "t_firstgen", "t_age", "n_course"),
                     p = TRUE, 
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_row_totals = TRUE,
                     include_row_missing_col = TRUE,
                     col_totals_rowname = "",
                     total_col = TRUE,
                     add_col_totals = TRUE,
                     col_totals_prefix = "N = ") %>%
  rename(N = "Total N", Missing = "Missing N")

@@

```{r racegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(8,4), fig.cap="Bar chart illustrating racial identification proprtions of participants"}

instructorrace <- expdata %>%
  dplyr::select(teacherid,
         t1_race_1, t1_race_2, t1_race_3, t1_race_4, t1_race_5, t1_race_6, t1_race_7, t_multi_race) %>%
  unique() %>%
  dplyr::summarize('White/Caucasian' = sum(t1_race_1, na.rm = TRUE)/n(),
            'Black or African American' = sum(t1_race_2, na.rm = TRUE)/n(),
            'Hispanic American or Latino/a' = sum(t1_race_3, na.rm = TRUE)/n(),
            'Asian or Pacific Islander' = sum(t1_race_4, na.rm = TRUE)/n(),
            'American Indian or Alaskan' = sum(t1_race_5, na.rm = TRUE)/n(),
            'Middle Eastern' = sum(t1_race_6, na.rm = TRUE)/n(),
            'Other Race' = sum(t1_race_7, na.rm = TRUE)/n(),
            'Multiple' = sum(t_multi_race > 1, na.rm = TRUE)/n())
studentrace <- expdata %>%
  dplyr::select(id,
         s1_race_1, s1_race_2, s1_race_3, s1_race_4, s1_race_5, s1_race_6, s1_race_7, s_multi_race) %>%
  unique() %>%
  dplyr::summarize('White/Caucasian' = sum(s1_race_1, na.rm = TRUE)/n(),
            'Black or African American' = sum(s1_race_2, na.rm = TRUE)/n(),
            'Hispanic American or Latino/a' = sum(s1_race_3, na.rm = TRUE)/n(),
            'Asian or Pacific Islander' = sum(s1_race_4, na.rm = TRUE)/n(),
            'American Indian or Alaskan' = sum(s1_race_5, na.rm = TRUE)/n(),
            'Middle Eastern' = sum(s1_race_6, na.rm = TRUE)/n(),
            'Other Race' = sum(s1_race_7, na.rm = TRUE)/n(),
            'Multiple' = sum(s_multi_race > 1, na.rm = TRUE)/n())

bind_cols(as_tibble(cbind(nms = names(instructorrace), t(instructorrace))),
          as_tibble(cbind(nms = names(studentrace), t(studentrace))))%>%
  dplyr::select(-nms...3) %>%
  pivot_longer(!nms...1, names_to = "group", values_to = "portion") %>%
  mutate(nms...1 = fct_relevel(nms...1, 
                               c('Hispanic American or Latino/a',
                                 'White/Caucasian',
                                 'Asian or Pacific Islander',
                                 'Middle Eastern',
                                 'Black or African American', 
                                 'American Indian or Alaskan',
                                 'Other Race',
                                 'Multiple')),
         group = fct_relevel(group, c("V2...4", "V2...2"))) %>%
  ggplot() +
  geom_col(aes(x = nms...1, y = as.numeric(portion), fill = group), position = "dodge") +
  labs(x = "Self-identified race/ethnicity", y = "Proportion of sample", 
       fill = "Group", title = "Self-identified racial proportions differ for instructors and students") +
  scale_fill_manual(values = c("#c8b5f7", "#f9b5c9"), labels = c("Students", "Instructors")) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 11)) +
  theme_minimal()
```

@@

```{r classsizegraph, echo=FALSE, warning=FALSE, message=FALSE, fig.dim=c(8.5,3), fig.cap="Histogram illustrating class sizes", out.width="100%", fig.align="center"}
n_instructor <- expdata %>%
  dplyr::select(teacherid, n_course) %>%
  unique()

@@

title <- ggdraw() + 
  draw_label(
    "Some classes are very large (class size means marked)",
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, 0, 7)
  )

plot_grid(title,
          combined,
          legend,
          ncol = 1,
          rel_heights = c(.09, 2))
```

@@

Table \@ref(tab:summarytable) shows summary statistics for key variables identified by @citeoriginal, including missing values. There do not appear to be significant differences between the treatment and control groups, other than in student similarity perception (p < 0.001 for initial perception, and p = 0.004 for end-of-term perception). Even in these two measures, the difference appears to be small (mean scores of 3.4 versus 3.6 for intial perception and 3.5 versus 3.6 for end-of-term perception).

@@

$X_{1i}$ is a vector of student-level covariates (student ISR anticipation, gender, CGPA, year of study).  The original study used gender and CGPA are used as controls for grade based outcomes only. @citeoriginal state this is because females earn higher grades generally. This paper uses them as controls for the majority of outcomes. ISR anticipation is used as a control only for end-of-term student ISR perception.

$X_{2j}$ is a vector of instructor-level covariates (class size). The original study consider instructor fixed effects instead. Because each participating instructor taught one class in the study, the class sizes correspond directly to the instructor IDs, so it was inadvisable to use both. This paper selected class size because of the larger potential for transferable insights.

@@

Equation \@ref(eq:ordlogmodel) is @citeoriginal's ordinal logistic model, used for the ordinal outcome, which is end-of-term instructor similarity perception. Because the instructors were only asked to answer one five-point scale question about similarity, this outcome cannot be treated the same way as the full-scale outcomes.

@@

Equation \@ref(eq:logmodel) is @citeoriginal's logistic model, used for the binary outcome, which is enrollment in Fall term 2017. Students are either not enrolled (0), enrolled (1), or they have graduated and are excluded from the model.

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Make all replication models
studentfirstsimilaritymodel <- lm(data = expdata,
                                  formula = s1_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
studentsecondsimilaritymodel <- lm(data = expdata,
                                   formula = s2_sim ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
studentsecondrelationshipmodel <- lm(data = expdata,
                                     formula = s2_tsr ~ treatment + s1_tsr + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
teachersecondrelationshipmodel <- lm(data = expdata,
                                     formula = t2_tsr ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
grademodel <- lm(data = expdata,
                 formula = grade ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
examgrademodel <- lm(data = expdata %>% filter(obj_exam == 1),
                     formula = t2_finalexam ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                     na.action = na.omit)
teachersecondsimilarity <- polr(formula = as.factor(t2_sim1) ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                                data = expdata,
                                Hess = TRUE,
                                     na.action = na.omit)
enrollmentfallmodel <- glm(data = expdata,
                           formula = f17_enrolled ~ treatment + n_course + s1_female + ir_f16_gpa + year,
                           family = "binomial",
                                     na.action = na.omit)
```

In Table \@ref(tab:modeltable), Model 1 shows that the treatment had a statistically significant effect on initial student similarity perception ($\beta$ = 0.194, 95% CI [0.136, 0.253]). Similarly, Model 2 that the treatment had a statistically significant effect on  student end-of-term similarity perception ($\beta$ = 0.11, 95% CI [0.038, 0.183]), though this effect size is smaller. Both coefficients are minimal on a scale of 1 to 5. The adjusted R-squared values, which indicate the proportion of variation in the outcome that can be explained by the treatment and covariates, are low (0.032 and 0.022). These models are helpful for analyzing treatment effect, but they do not aim to be and are not useful predictors of student similarity perception, not even within the sample.

Model 3 shows no evidence that treatment impacted student end-of-term ISR perception. However, it does show a significant association between student ISR anticipation and student end-of-term ISR perception ($\beta$ = 0.671, 95% CI [0.624, 0.717]). For every 1 point increase in ISR anticipation on a 1 to 5 scale, the expected increase in end-of-term ISR perception is 0.671. Models 4, 5, 6, 7, and 8 show no significant treatment effect on end-of-term instructor ISR perception, course grade, objectively graded exam grade, end-of-term instructor similarity perception, or enrollment in the subsequent term.

Model performance and validation for the two models with significant treatment effects, Models 1 and 2 from Table \@ref(tab:modeltable), can be found in [Appendix D].

```{r modeltable, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
# Make table to display model results
stargazer(studentfirstsimilaritymodel,
          studentsecondsimilaritymodel,
          studentsecondrelationshipmodel,
          teachersecondrelationshipmodel,
          grademodel,
          examgrademodel,
          teachersecondsimilarity,
          enrollmentfallmodel,
          title = "Replication model results",
          align = TRUE,
          ci = TRUE,
          float.env = "sidewaystable",
          notes.label = "",
          font.size = "small",
          no.space = TRUE,
          keep.stat = c("n", "rsq", "adj.rsq", "aic"),
          column.sep.width = "-10pt",
          label = "tab:modeltable",
          type = "latex",
          header = FALSE,
          initial.zero = FALSE,
          #omit = c("teacherid"),
          #omit.labels = c("Instructor fixed effect"),
          dep.var.labels = c("Student sim. 1",
                             "Student sim. 2",
                             "Student ISR 2",
                             "Instructor ISR 2",
                             "Course grade",
                             "Final grade",
                             "Instructor sim. 2",
                             "Persistence"),
          covariate.labels = c("Treatment", "Student ISR 1", "Class size", "Student female", "CGPA", "Year of study"))
```

@@

Similarly, the subset of first-generation college student participants is considered. The effect of treatment on student similarity perception is larger than for the general student population and the Black and/or Hispanic student participants ($\beta$ = 0.265, 95% CI [0.177, 0.353] for initial perception; $\beta$ = 0.15, 95% CI [0.039, 0.26] for end-of-term perception), but there is still no effect on ISR perception, instructor perception, or student outcomes. Full results are in [Appendix F].
